---
title: "IID Gamma Simulation"
format: html
---

```{r}
library(dplyr)
library(purrr)
library(tibble)
library(tidyr)
library(ggplot2)
library(knitr)
```

# IID Gamma simulation

This report summarizes the Monte Carlo simulation study for the IID Gamma goodness-of-fit test implemented in `gofedf::testGamma()`.

Data are generated from Gamma distributions with scale equal to 1 and shape parameters 1, 7, and 50. For each setting, we examine the distribution of p-values and estimate the empirical type-I error rate at the nominal level 0.05.

```{r}
sample_sizes <- c(25, 50, 100, 250)
shape_values <- c(1, 7, 50)
discretize_values <- c(FALSE, TRUE)
alpha <- 0.05
```

## Read p-values

```{r}
read_gamma_pvalues <- function(n, shape, discretize) {

  file <- paste0(
    "../results/gamma/gamma_iid_n", n,
    "_shape", shape,
    "_discretize_", discretize,
    "_pvalues.rds"
  )

  if (!file.exists(file)) {
    stop("File does not exist: ", file)
  }

  pvals <- readRDS(file)

  tibble(
    n = n,
    shape = shape,
    scale = 1,
    discretize = discretize,
    rep = seq_along(pvals),
    cvm_pvalue = map_dbl(pvals, 1),
    ad_pvalue = map_dbl(pvals, 2)
  )
}
```

```{r}
gamma_results <- expand.grid(
  n = sample_sizes,
  shape = shape_values,
  discretize = discretize_values
) |>
  pmap_dfr(function(n, shape, discretize) {
    read_gamma_pvalues(
      n = n,
      shape = shape,
      discretize = discretize
    )
  })

gamma_long <- gamma_results |>
  pivot_longer(
    cols = c(cvm_pvalue, ad_pvalue),
    names_to = "method",
    values_to = "pvalue"
  ) |>
  mutate(
    method = recode(
      method,
      cvm_pvalue = "CvM",
      ad_pvalue = "AD"
    )
  )
```

## Histogram of p-values

### CvM p-values

```{r}
gamma_long |>
  filter(method == "CvM") |>
  ggplot(aes(x = pvalue)) +
  geom_histogram(bins = 30, boundary = 0) +
  facet_grid(shape + discretize ~ n) +
  labs(
    x = "p-value",
    y = "Count",
    title = "Histogram of CvM p-values: IID Gamma simulation"
  ) +
  theme_minimal()
```

### AD p-values

```{r}
gamma_long |>
  filter(method == "AD") |>
  ggplot(aes(x = pvalue)) +
  geom_histogram(bins = 30, boundary = 0) +
  facet_grid(shape + discretize ~ n) +
  labs(
    x = "p-value",
    y = "Count",
    title = "Histogram of AD p-values: IID Gamma simulation"
  ) +
  theme_minimal()
```

## Type-I error rate

```{r}
type_one_table <- gamma_long |>
  group_by(n, shape, scale, method, discretize) |>
  summarise(
    MC = n(),
    type_one_error = mean(pvalue < alpha, na.rm = TRUE),
    .groups = "drop"
  )
```

### CvM type-I error rate

```{r}
type_one_table |>
  filter(method == "CvM") |>
  arrange(shape, discretize, n) |>
  kable(
    digits = 4,
    caption = "Estimated type-I error rates for CvM: IID Gamma simulation"
  )
```

### AD type-I error rate

```{r}
type_one_table |>
  filter(method == "AD") |>
  arrange(shape, discretize, n) |>
  kable(
    digits = 4,
    caption = "Estimated type-I error rates for AD: IID Gamma simulation"
  )
```
